Stochastic Convolutional Sparse Coding

### 2018-10-30 cvpr2019/Conclusion.tex
minor edits

\section{Conclusions}
In this work, we present a novel stochastic subsampling strategy for solving the CSC problem in spatial domain. This method significantly improves the runtime performance over the prior frequency-domain solvers, which applies to both batch mode and online-learning mode, as well as generating effective and proven outcomes. The proposed algorithm, for the first time, demonstrates the feasibility that tackling the CSC problem in spatial domain while still holding the runtime efficiency. Since the subproblem of updating the code is a highly sparse LASSO, other specific optimization strategies can be applied to further accelerate the computation, for instance the idea proposed in~\cite{johnson2017stingycd}, which solves the LASSO problem by coordinate descent and skips unnecessary updates by the method of safe screening~\cite{ghaoui2012Swfe}.

Furthermore, instead of employing uniform subsampling strategy, alternative subsampling strategies can be studied in the future for a gain of learning efficiency, for instance an importance or correlation based sampling strategy which considers the dependency relationship between the images and the sparse codes. Notice that transforming the problem into Fourier domain cannot benefit from these kinds of speedup strategies (including the proposed one).

>>>>

\section{Conclusions}
In this work, we present a novel stochastic subsampling strategy for solving the CSC problem in spatial domain. This method significantly improves the runtime performance over the prior frequency-domain solvers, which applies to both batch mode and online-learning mode, as well as generating effective and proven outcomes. The proposed algorithm, for the first time, demonstrates the feasibility that tackling the CSC problem in spatial domain while still holding, or even improving the runtime efficiency. Since the subproblem of updating the code is a highly sparse LASSO, other specific optimization strategies can be applied to further accelerate the computation, for instance the idea proposed in~\cite{johnson2017stingycd}, which solves the LASSO problem by coordinate descent and skips unnecessary updates using the method of safe screening~\cite{ghaoui2012Swfe}. It would be worthy to emphasize that transforming the problem into Fourier domain cannot benefit from these kinds of speedup strategies (including the proposed one). Furthermore, instead of employing uniform subsampling strategy, alternative subsampling strategies can be studied in the future for a gain of learning efficiency, for instance an importance or correlation based sampling strategy which considers the dependency relationship between the signals and the sparse codes.

### 2018-10-30 cvpr2019/Results.tex
minor edits

{\bfseries Convergence.} Comparisons of the convergence between the proposed method (SBCSC) and the state-of-the-art batch-mode algorithm~\cite{heide2015fast} are shown in Fig.\ \ref{fig:subsampleResult}. A different selection of the subsampling rate reveals that the proposed strategy will slightly influence the convergence and the training objective of the minimization problem. Specifically, the more subsampled, the relatively slower convergence and the higher objective will be obtained. On the other hand, small subsampling rate will significantly accelerate the computation process, where $10\%$ subsampling achieves about $6 \times$ speedup over no subsampled solver and $2 \times$ speedup over state-of-the-art frequency solver for one iteration. We observe that a subsampling rate between $0.1$ and $0.2$ delivers empirically good enough convergence in our settings, as well as achieving at least $3 \times$ speedup. In general, the proposed method with various subsampling rates converges at around 10-12 iterations in all testing cases, acting similar to the competing methods. Notice that the comparable method has a relatively high objective at first $6-8$ iterations due to different splitting strategies are applied and subproblems are interleaved on the primary variables~\cite{wohlberg2016efficient}, though it converges to an optimum with comparable objective after 10 iterations. In summary, the convergence behaviors of the proposed algorithm is only slightly influenced by the subsampling strategy within the testing subsampling rates. Comparing to the state-of-the-art frequency solver, the proposed stochastic spatial-domain solver with a subsampling rate of $0.1$ reduces the computation time by a factor of two for the tested example. Please see supplementary materials for additional experimental results.

@@

{\bfseries Convergence.} Unlike the batch-based learning approaches which evaluate its convergence by monitoring the objective value on training datasets, a common way to evaluate the learning process of online learning model is to monitor its objective value on testing datasets. The testing objectives of the proposed method (SOCSC) and prior published online frequency-domain CSC method~\cite{liu-2018-first}, with respect to iterations, are shown in Fig.\ \ref{fig:onlineSmall}. In the same figure, we also keep track of the capability of the updated filters during the learning process to sparsely represent the testing images, which is demonstrated by the time evolution of PSNR. These two approaches stop at optimum positions with close testing objective values, though, they exhibit slightly different convergence behaviors with respect to iterations. The final testing PSNR values for both methods also reveal a similar reconstruction performance of the learned filters. As for the runtime comparison, the proposed method runs at least 5 times faster than the comparable method. Supplement shows additional comparisons for the other dataset.

>>>>

{\bfseries Convergence.} Comparisons of the convergence between the proposed method (SBCSC) and the state-of-the-art batch-mode algorithm~\cite{heide2015fast} are shown in Fig.\ \ref{fig:subsampleResult}. A different selection of the subsampling rate reveals that the proposed strategy will slightly influence the convergence and the training objective of the minimization problem. Specifically, the more subsampled, the relatively slower convergence and the higher objective will be obtained. On the other hand, small subsampling rate will significantly accelerate the computation process, where $10\%$ subsampling achieves about $6 \times$ speedup over no subsampled solver and $2 \times$ speedup over state-of-the-art frequency solver for one iteration. We observe that a subsampling rate between $0.1$ and $0.2$ delivers empirically good enough convergence in our settings, as well as achieving at least $3 \times$ speedup. In general, the proposed method with various subsampling rates converges at around 10-12 iterations in all testing cases, acting similar to the competing methods. Notice that the comparable method has a relatively high objective at first $6-8$ iterations due to different splitting strategies are applied and subproblems are interleaved on the primary variables~\cite{wohlberg2016efficient}, though it converges to an optimum with comparable objective after 10 iterations. In summary, the convergence behaviors of the proposed algorithm is only slightly influenced by the subsampling strategy within the testing subsampling rates. Comparing to the state-of-the-art frequency solver, the proposed stochastic spatial-domain solver with a subsampling rate of $0.1$ reduces the computation time by a factor of two for the tested example. The robustness of the proposed algorithm is evaluated by additional experiments. Please see supplementary materials for reference.

@@

{\bfseries Convergence.} Unlike the batch-based learning approaches which evaluate its convergence by monitoring the objective value on training datasets, a common way to evaluate the learning process of online learning model is to monitor its objective value on testing datasets. The testing objectives of the proposed method (SOCSC) and prior published online frequency-domain CSC method~\cite{liu-2018-first}, with respect to iterations, are shown in Fig.\ \ref{fig:onlineSmall}. In the same figure, we also keep track of the capability of the updated filters during the learning process to sparsely represent the testing images, which is demonstrated by the time evolution of PSNR. These two approaches stop at optimum positions with close testing objective values, though, they exhibit slightly different convergence behaviors with respect to iterations. The final testing PSNR values for both methods also reveal a similar reconstruction performance of the learned filters. As for the runtime comparison, the proposed method runs at least 5 times faster than the comparable method. Supplement shows additional comparisons for the other datasets.

### 2018-10-30 cvpr2019/SCSCmodel.tex
minor edits

The work of~\cite{mensch2016dictionary} also applies a so-called stochastic subsampling strategy for dictionary learning, while the subsampling is randomly performed on the observed data, differing from that of our proposed. One can think of our proposed randomization approach as a variant of the stochastic Frank-Wolfe algorithm~\cite{reddi2016stochastic,pmlr-v80-kerdreux18a}, in which the optimization are performed on a subset of the variables which are randomly extracted based on a certain probability distribution at each iteration. For the proposed algorithm, each iteration extracts ($pDK+MK$) variables, where $pDK$ variables are randomly picked out from $DK$ variables with identical probability and the rest remain unchanged. Owing to the fact that the codes are highly sparse, the original signals can still be represented by a portion of the codes with a reasonable subsampling rate. Herein, unlike the stochastic Frank-Wolfe algorithm that generally requires more iterations to reach convergence, the convergence of the proposed algorithm will not be significantly affected by the subsampling manipulation. This will be experimentally verified in Section~\ref{sec:result}.

>>>>

The work of~\cite{mensch2016dictionary} also applies a so-called stochastic subsampling strategy for dictionary learning, while the subsampling is randomly performed on the
observed signals, differing from that of our proposed. In addition, it is basically applied for factorizing matrix that is large in both dimensions. Even though CSC can be 
expressed in a similar mathematical formula as matrix factorization, they are essential different due to the convolution properties in CSC model.
One can think of our proposed randomization approach as a variant of the stochastic Frank-Wolfe algorithm~\cite{reddi2016stochastic,pmlr-v80-kerdreux18a}, in which the optimization are performed on a subset of the variables which are randomly extracted based on a certain probability distribution at each iteration. For the proposed algorithm, each iteration extracts ($pDK+MK$) variables, where $pDK$ variables are randomly picked out from $DK$ variables with identical probability and the rest remain unchanged. Owing to the fact that the codes are highly sparse, the original signals can still be represented by a portion of the codes with a reasonable subsampling rate. Herein, unlike the stochastic Frank-Wolfe algorithm that generally requires more iterations to reach convergence, the convergence of the proposed algorithm will not be significantly affected by the subsampling manipulation. This will be experimentally verified in Section~\ref{sec:result}.

### 2018-10-30 cvpr2019/main.tex
minor edits

%%%%%%%%% ABSTRACT
\begin{abstract}
The minimization problems formulated in Convolutional Sparse Coding (CSC) have been solved in frequency domain by modern approaches owing to the improved computing efficiency. However, solving the problem in frequency domain automatically assumes circular boundary conditions, and also it is restricted by imposing modern optimization strategies for this specific problem. In this work, we propose a novel stochastic spatial-domain solver where a randomized subsampling strategy is introduced in the subproblem of computing sparse codes. Afterwards, we extend the proposed strategy in conjunction with online learning, scaling the CSC model up to arbitrary sample sizes. In both cases, we show experimentally that the proposed method outperforms the state-of-the-art frequency-domain solvers with a reasonable selection of the subsampling rate. At last, we learn the over-complete dictionary from large-scale datasets, which demonstrates a better sparse representation of the natural images on account of more abundant learned image features.
\end{abstract}

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
The minimization problems formulated in Convolutional Sparse Coding (CSC) have been solved in frequency domain by modern approaches owing to the improved computing efficiency. However, solving the problem in frequency domain automatically assumes circular boundary conditions, and also it is restricted by imposing modern optimization strategies for this specific problem. In this work, we propose a novel stochastic spatial-domain solver, in which a randomized subsampling strategy is introduced in the subproblem of computing sparse codes. Afterwards, we extend the proposed strategy in conjunction with online learning, scaling the CSC model up to arbitrary sample sizes. In both cases, we show experimentally that the proposed method, with a reasonable selection of the subsampling rate, outperforms the state-of-the-art frequency-domain solvers in terms of execution time without losing the learning quality. At last, we evaluate the effectiveness of the over-complete dictionary learned from large-scale datasets that has not been studied by precedent CSC work, which demonstrates an improved sparse representation of the natural images on account of more abundant learned image features.
\end{abstract}
